---
title: "P8105 Homework 5"
author: "Courtney Diamond"
date: "2023-11-13"
output: github_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(magrittr)
library(ggridges)
library(lubridate)


knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

First up, we are going to load the specified data. I have downloaded the CSV from the github and saved it into a `data` file within this repository.

```{r}
homicide_data = read_csv("data/homicide-data.csv")
```

The raw data consists of `r ncol(homicide_data)` variables and `r nrow(homicide_data)` individual observations. The variables are: 

* `uid`: a unique identifier for each homicide
* `reported_date`: the date the homicide was reported, in yyyymmdd format
* `victim_last`: the victim's last name
* `victim_first`: the victim's first name
* `victim_race`: the victim's race
* `victim_age`: the victim's age
* `victim_sex`: the victim's sex (male, female, or unknown)
* `city`: the city of the homicide
* `state`: the state of the homicide
* `lat`: the lattitude of the homicide
* `long`: the longitude of the homicide
* `disposition`: the case status, i.e. 'Closed by arrest', 'Closed without arrest', or 'Open/No arrest'



```{r}
solved_unsolved_ratio_table =
  homicide_data |> 
  mutate(city_state = str_c(as.character(city), as.character(state), sep = ", ")) |> 
  filter(city_state != 'Tulsa, AL') |> 
  mutate(status = case_when(
    disposition == "Closed by arrest" ~ "solved",
    disposition == "Closed without arrest" ~ "unsolved",
    disposition == "Open/No arrest" ~ "unsolved"
  )) |>
  group_by(city_state) |> 
  summarize(all_homicides = n(),
            unsolved_homicides = sum(status == "unsolved"))

solved_unsolved_ratio_table

solved_unsolved_ratio_table |> 
  knitr::kable()
  
```

```{r}
baltimore_homicide_prop = 
  prop.test(x = filter(solved_unsolved_ratio_table, city_state == 'Baltimore, MD') |>
              pull(unsolved_homicides),
            n = filter(solved_unsolved_ratio_table, city_state == 'Baltimore, MD') |>
              pull(all_homicides)) |> 
  broom::tidy() |> 
  select(estimate, conf.low, conf.high)

baltimore_homicide_prop
```

Ok, we worked out an example case for baltimore. Let's make map these functions to the whole dataframe.  
```{r}
calculate_unsolved_prob =
  solved_unsolved_ratio_table |> 
  mutate(prop_results = 
           map2(unsolved_homicides, all_homicides, \(x, y) prop.test(x = x, n = y)),
         tidy_results = 
           map(prop_results, broom::tidy)) |> 
  select(city_state, tidy_results) |> 
  unnest(tidy_results) |> 
  select(city_state, estimate, conf.low, conf.high) |> 
  mutate(city_state = fct_reorder(city_state, estimate))
  

calculate_unsolved_prob

calculate_unsolved_prob |> 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  geom_point() +
  theme(axis.text.x  = element_text(angle = 45, vjust = 1, hjust = 1))


```

## Problem 2
